Mapping Cancer: the Potential of Cartograms and Alternative Map Displays

### 2020-05-17 literature.Rmd
Adjusted heading levels

# Traditional approaches for cancer map displays {#ch:choropleth-maps}

@@

# Contemporary alternatives to choropleth maps {#ch:alternatives}

## Cartograms
<!-- 1496 words inc table-->

@@

## Tile Map
<!-- 173 words -->
A tile map provides a tessellated display of consistent shapes.  A similar method to a rectangular cartogram, represents each geographic area using a square. The squares are tessellated to create a grid. Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile. Regions with over four neighbors require some necessary displacement. The tile map uses color to represent a value of a statistic for each area. A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions. There are online media sources using this method, these include [@NPR], [@FiveThirtyEight], [@WSJ], [@WP]. Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.

@@

## Geofacet
<!-- 166 words -->
Hafen [@IGF] introduces the term geofacet to describe a grid display of small plots. The arrangement of tiles mimics the geographic topology. Geofaceting has the functionality that a statistical plot can be constructed in each facet for each geographic area. A tile map can communicate only one value per region in a visualization, while geofaceting is a more flexible visualization for communication as it increases the amount of information displayed. Virtually any type of plot can be shown in the tile, allowing displays of multiple variables or values per geographic entity. Creating the layout of a geofacet is manual, but once created can be used for any data on that geographic base.

@@

## Multivariate displays
<!-- 222 words -->

@@

## Neither choropleth maps or cartograms perform well for Australia

>>>>

## Traditional approaches for cancer map displays {#ch:choropleth-maps}

@@

## Contemporary alternatives to choropleth maps {#ch:alternatives}

### Cartograms
<!-- 1496 words inc table-->

@@

### Tile Map
<!-- 173 words -->
A tile map provides a tessellated display of consistent shapes.  A similar method to a rectangular cartogram, represents each geographic area using a square. The squares are tessellated to create a grid. Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile. Regions with over four neighbors require some necessary displacement. The tile map uses color to represent a value of a statistic for each area. A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions. There are online media sources using this method, these include [@NPR], [@FiveThirtyEight], [@WSJ], [@WP]. Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.

@@

### Geofacet
<!-- 166 words -->
Hafen [@IGF] introduces the term geofacet to describe a grid display of small plots. The arrangement of tiles mimics the geographic topology. Geofaceting has the functionality that a statistical plot can be constructed in each facet for each geographic area. A tile map can communicate only one value per region in a visualization, while geofaceting is a more flexible visualization for communication as it increases the amount of information displayed. Virtually any type of plot can be shown in the tile, allowing displays of multiple variables or values per geographic entity. Creating the layout of a geofacet is manual, but once created can be used for any data on that geographic base.

@@

### Multivariate displays
<!-- 222 words -->

@@

## Performance of choropleth maps or cartograms for Australia
